\vskip-0.6in Trade Agreements in the Shadow of Lobbying\protect\footnote{Formerly circulated under the title ``Trade Agreements, Lobbying and Separation of Powers''}

### 2016-02-02 draft.tex
capitalized a reference title
write two new opening paragraphs for intro

Empirical investigations of Grossman and Helpman's (1994) `Protection for Sale' model (henceforth `PFS') broadly support the predictions of the theory qualitatively, yet their quantitative estimates present several interesting puzzles.\footnote{cfr. Goldberg and Maggi (1999); Gawande and Bandyopadhyay (2000); Mitra, Thomakos and Ulubasoglu (2002); McCalman (2004).\label{fn:gh_lobby}} These studies have consistently found the weight governments place on social welfare to be many times that which they place on lobbying effort, while estimates indicate that the deadweight losses caused by trade distortions are several orders of magnitude larger than lobbying expenditures.\footnote{Feenstra (1992) assembles estimates from the mid-1980s that place a floor of $\$$8 billion per year on the deadweight losses from protection, whereas Bombardini and Trebbi (2012) find that total annual lobbying expenditure on trade issues in 1999-2001 when this data first became available was about $\$$200 million.} This raises the question: how could governments value social welfare so highly, yet grant these quantities of protection at such a low price?

@@

Since so much of trade policy is determined in the context of international agreements that must be ratified by domestic legislative bodies, I incorporate PFS-style lobbying into a model of trade agreements in which the executive branches of the governments set tariff levels in anticipation of political pressure upon the legislatures. This political pressure may be interpreted either as lobbying against ratification of the trade agreement or lobbying to break the agreement once in force. I find that the separation-of-powers government structure can explain the empirical puzzle surrounding governments' welfare weights and that the addition of political uncertainty adds important realism so that the model presented here can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.

@@

\item Le Breton, M., Salanie, F., 2003. Lobbying under political uncertainty. Journal of Public Economics 87, 2589-2610.

>>>>

Much of trade policy is determined in the context of international agreements that must be ratified by some domestic body. This paper seeks to shed light on such a policy-making process by incorporating Protection-for-Sale (Grossman and Helpman 1994, hereafter PFS) style lobbying into a model of trade agreements in which the executive branches of the governments set tariff levels in anticipation of political pressure upon a ratifying body, assumed here to be a legislature. This political pressure may be interpreted either as lobbying against ratification of the trade agreement or lobbying to break the agreement once in force. Important realism is achieved not only through the addition of the ratification constraint but also by the incorporation of political uncertainty into the model.

Assuming for now that lobbies represent only import-competing interests, lobbies in this environment would like to prevent or derail a trade agreement that would reduce their tariff protection. They work toward this end by exerting effort to convince their legislatures to vote against the trade agreement but face some uncertainty about whether their efforts will result in a voting outcome consistent with their interests. I show that the probability that the trade agreement is derailed increases in lobbying effort and decreases in the trade agreement tariff. Further, the lobby decreases its effort as the tariff it would face under the trade agreement rises: it has less to lose the higher is the trade agreement tariff. When forming the trade agreement tariff, the executives trade off the welfare they receive under the agreement with the probability that the agreement actually remains in force. Ultimately, I find that observed lobbying effort may often be low because the executive branches of the governments set trade agreement tariffs sufficiently high to reduce political pressure on the legislatures.

\vskip.2in
I find that the separation-of-powers government structure can explain the empirical puzzle surrounding governments' welfare weights and that the addition of political uncertainty adds important realism so that the model presented here can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.

Empirical investigations of Grossman and Helpman's (1994) PFS model broadly support the predictions of the theory qualitatively, yet their quantitative estimates present several interesting puzzles.\footnote{cfr. Goldberg and Maggi (1999); Gawande and Bandyopadhyay (2000); Mitra, Thomakos and Ulubasoglu (2002); McCalman (2004).\label{fn:gh_lobby}} These studies have consistently found the weight governments place on social welfare to be many times that which they place on lobbying effort, while estimates indicate that the deadweight losses caused by trade distortions are several orders of magnitude larger than lobbying expenditures.\footnote{Feenstra (1992) assembles estimates from the mid-1980s that place a floor of $\$$8 billion per year on the deadweight losses from protection, whereas Bombardini and Trebbi (2012) find that total annual lobbying expenditure on trade issues in 1999-2001 when this data first became available was about $\$$200 million.} This raises the question: how could governments value social welfare so highly, yet grant these quantities of protection at such a low price?

@@

@@

\item Le Breton, M., Salanie, F., 2003. Lobbying under Political Uncertainty. Journal of Public Economics 87, 2589-2610.
